---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# poemReport


The goal of poemReport is to make reporting of statistical services automatic.

## Installation

You can install poemReport using

``` r
install.packages("devtools")
devtools::install_github("dpananos/poemReport")
```

## Reporting

To begin the delays analysis for a given month, use `poemReport::setup()`:

```{r, eval=FALSE}
# Insert the appropirate month and year of the report
poemReport::setup(report_month = "May", 
                    report_year = "2020")
```

This will open a `.R` file which looks like

```{r}
# Create POEM Report for May 2020


poemReport::data_cleaning(report_month = "May", report_year = "2020")

poemReport::poem_report(report_month = "May", report_year = "2020")

```


Running `poemReport::data_cleaning(report_month = "May", report_year = "2020")` opens a cleaning script called `02-data_cleaning.R`.  Analysts are required to point the script towards the appropriate databases in order to analyze them.  The remainder of the script implements analyses for statistical services data.  The analyst is free to add/remove/change any of the analyses as they see fit.

Running `02-data_cleaning.R` will save a copy of the clean data in a `.RDS` file into the working directory.  The report can be created by running `poemReport::poem_report(report_month = "May", report_year = "2020")` and knitting the document.
