---
title: "POEM Statistical Services Report"
author: ""
date: "{{{ report_month }}} {{{ report_year }}}"
output: 
  pdf_document:
    latex_engine: xelatex
mainfont: Arial
---


```{r, include=F}
#These are some options for rendering to word
knitr::opts_chunk$set(
  echo = F, #Do not print the code
  message = F, #Do not show error messages
  warning = F, #Do not show warnings
  fig.height = 3, #Height of images in inches
  fig.width = 5, #Width of images in inches
  fig.align = 'center', #Center the images
  dpi = 400, #Image quality
  cache = F #Cache computations from previous renders
)
```


```{r, load_data}
library(kableExtra)
library(knitr)
library(lubridate)
library(stringr)
library(tidyverse)

theme_set(theme_bw())

#Read in the data to be analyzed
work <-
  readRDS(
    here::here(
      "reports",
      "{{{ report_year }}}",
      "{{{ report_month }}}",
      "delays_{{{ report_month }}}_{{{ report_year }}}_clean.rds"
    )
  )

current.year<- as.numeric({{{ report_year }}})
```

# Service Usage

```{r, service_access_count}
#How many times were we asked for service
#Equivalent to asking: how many unique projects are there?
#Unique request determined byh combination of Name, type, and ID
num_access_serv = work %>%
  arrange(WorkDate) %>%
  distinct(Name, type, ID) %>%
  group_by(type) %>%
  summarise(n = n())

#How many unique people made these requests
num_unique_memb = work %>% distinct(Name) %>% nrow()
```

```{r, hours_used}
#Summarize how many hours were used per member
#Roll up to determine usage by year
work_hrs_crrnt_yr <- work %>%
  mutate(Year = year(WorkDate)) %>%
  group_by(Name, type, Year) %>%
  summarise(hrs = sum(WorkHrs)) %>%
  filter(Year == current.year) %>%
  ungroup

#Summary of work hours is given on a per member basis
#For instance, must report the median number of hours used per member.
smmry_wrk_hrs <- work_hrs_crrnt_yr %>%
  summarise(min = min(hrs),
            max = max(hrs),
            med = median(hrs))
```


```{r, grant_reviews}
#Count how many grant reviews we did for each agency
#Need to break up the GrantAgency Column since it may contain multiple grant agencies
#Count up AMOSO and CIHR.  Everything else is under OTHER
grt_rvw_cnt <- work %>%
  filter(type == 'GRANT') %>%
  arrange(WorkDate) %>%
  distinct(Name, ID, type, .keep_all = T) %>%
  transmute(agency = str_split(GrantAgency, ';')) %>%
  mutate(cihr = map_int(agency, ~ any(c('CIHR') %in% .x)),
         amoso = map_int(agency, ~ any(c('AMOSO') %in% .x)),
  ) %>%
  summarise_at(vars(cihr, amoso), function(x) sum(x, na.rm = T))
 
```

* A total of `r sum(num_access_serv$n)` requests for service have been made since the services inception.  This includes requests for statistical services as well as grant reviews. These requests have been made by  `r num_unique_memb` unique POEM members.

* A total of `r sum(work_hrs_crrnt_yr$hrs)` hours have been used per DOM member in `r current.year` (median `r smmry_wrk_hrs$med` hours per DOM faculty member ), with the number of hours used ranging from `r smmry_wrk_hrs$min` to `r smmry_wrk_hrs$max` per project.

* `r grt_rvw_cnt$cihr` CIHR review(s) have been conducted. `r grt_rvw_cnt$amoso` AMOSO grant review(s) have been conducted. 

# Requested Services

Shown below is a table of requested services and the proportion of projects which required the service.  Note that the proportions do not sum to 100% because some projects require multiple services (note, service began July 2018).


```{r, work_request_summary }


work_rqst<-work %>% 
  mutate(Year = year(WorkDate)) %>% 
  arrange(WorkDate) %>% 
  distinct(Name, type, ID, .keep_all = T) %>% 
  mutate(work_request = str_split(WorkRequested,';')) %>% 
  unnest() %>% 
  mutate(work_request = str_to_title(work_request),
         work_request = str_trim(work_request)) %>% 
  group_by(Year, work_request) %>% 
  summarise(n = n()) %>% 
  mutate(freq = n/sum(n)) %>% 
  select(-n) %>% 
  mutate(freq = scales::percent(freq)) %>% 
  spread(Year, freq, fill = '0%') %>% 
  rename(`Work Request` = work_request)
  
#make table
kable(work_rqst, 
      format = 'latex', 
      align = rep('r', ncol(work_rqst)),
      booktabs = T,
      caption = 'Proportion of requests in indicated year which requested indicated service.  Sums may exceed 100\\% as service requests can request several services.') %>% 
  kable_styling(latex_options = c('striped',"condensed", 'hold_position'), full_width = F)
  
```

# Service Demand

Shown below is a plot of service demand since service has been introduced.  Service is expected to fluctuate as various deadlines approach.

```{r}
demand <- 
  work %>% 
  arrange(WorkDate) %>% 
  distinct(Name, ID, type, .keep_all = T) %>% 
  mutate(Time = date(floor_date(WorkDate, 'month'))) %>%  
  drop_na(Time) %>% 
  count(Time) %>% 
  complete(Time = seq.Date(min(Time),max(Time), by = 'month')) %>% 
  replace_na(list(n=0))

demand %>% 
  ggplot(aes(Time,n))+
  geom_col(fill = 'light gray', color = 'black')+
  scale_y_continuous(limits = c(0,NA))+
  scale_x_date(date_breaks = '3 months', date_labels = '%Y/%m')+
  labs(x = 'Date', y = 'Requests in Month')+
  theme(plot.margin=unit(c(1,1,1,1),"cm"))
```

  
# Service Adoption

```{r, unique_members}

growth <-
  work %>%
  arrange(WorkDate) %>%
  distinct(Name, .keep_all = T) %>%
  mutate(qtr = stringr::str_c(year(WorkDate), '-Q', quarter(WorkDate)),
         QTR = factor(qtr)) %>%
  group_by(QTR) %>%
  summarise(n = n_distinct(Name)) %>%
  ungroup %>%
  mutate(n = cumsum(n))


growth.rate <- mean(diff(growth$n))

```

New clients adopting the service is an important metric to monitor.  Shown below is a plot of total unique members over time.  To date, the service has seen an adoption rate of `r growth.rate` DOM members per quarter.  That is to say `r growth.rate` DOM members use the service for the first time per quarter on average.

```{r, service_growth}

growth %>%
  ggplot(aes(QTR, n)) +
  geom_col(fill = 'light gray', color = 'black') +
  geom_label(aes(label = n)) +
  labs(x = 'Year', y = 'Cumulative DOM Members\nUsing Service')
```



# Clients By Department
## Including SAS requests, Grant Reviews, and Stats Related REB Protocol Reviews Since Program Inception


```{r}
#Create table for division counts stratified by DOM LEVEL
division.requests <- 
  work %>%
  filter(type == 'SAS') %>%
  arrange(WorkDate) %>%
  distinct(Name, ID, .keep_all = T) %>%
  select(Division, DOMLevel) %>%
  group_by(Division, DOMLevel) %>%
  count() %>%
  replace_na(list(DOMLevel = 'Not Listed')) %>%
  ungroup

#Tack on how many trainees were added
trainees <- 
  work %>%
  filter(type == 'SAS') %>%
  arrange(WorkDate) %>%
  distinct(Name, ID, .keep_all = T) %>%
  select(Division, Collaborator) %>%
  mutate(collab_list = map_int(stringr::str_split(Collaborator, ';'), length)) %>%
  group_by(Division) %>%
  summarise(Trainees = sum(collab_list)) %>%
  ungroup


division.requests.table <- 
  division.requests %>%
  spread(DOMLevel, n) %>%
  replace_na(list(
    JR = 0,
    SR = 0,
    MID = 0,
    `Not Listed` = 0
  )) %>%
  mutate(SAS = JR + SR + MID + `Not Listed`) %>%
  arrange(desc(SAS)) %>%
  select(Division, SAS, JR, MID, SR, `Not Listed`) %>%
  left_join(trainees) %>%
  replace_na(list(trainees = 0))

requests.total <- 
  division.requests.table %>%
  select(-Division) %>%
  summarise_all(list(sum)) %>%
  mutate(Division = 'Total')

requests<-
division.requests.table %>%
bind_rows(requests.total)

kable(requests,
      format = 'latex', 
      align = rep('r', ncol(requests)),
      booktabs = T,
      caption = 'Requests for statistical services stratified by division and seniority.') %>% 
  kable_styling(latex_options = c('striped',"condensed", 'hold_position'), full_width = F)
  
```



```{r}
repeat.clients <-
  work %>%
  distinct(Name, ID, type, .keep_all = T) %>%
  group_by(Name, DOMLevel) %>%
  count() %>%
  arrange(desc(n)) %>%
  filter(n > 1)

num.mids <- filter(repeat.clients, DOMLevel == 'MID') %>% nrow
num.jr <- filter(repeat.clients, DOMLevel == 'JR') %>% nrow
num.sr <- filter(repeat.clients, DOMLevel == 'SR') %>% nrow

num.residents <- 
  work %>%
  distinct(Name, ID, .keep_all = T) %>%
  drop_na(PGYNom) %>%
  mutate(pgy = map_int(stringr::str_split(PGYNom, ';'), length)) %>%
  summarise(pgy = sum(pgy))


```

We have had `r nrow(repeat.clients)` repeat client(s) (`r num.jr` JR, `r num.mids` MID, `r num.sr` SR).  A new SAS form is not required for continuing work on the same study.  We encourage trainees -- post docs, RA/RC -- to attend stats meetings.  A total of `r num.residents$pgy` resident projects were helped by statistical services.


## Grant Reviews Since Program Inception

```{r}
grant.reviews <- 
  work %>%
  filter(type == 'GRANT') %>%
  arrange(WorkDate) %>%
  distinct(type, ID, .keep_all = T) %>%
  select(Division, DOMLevel, GrantAgency) %>%
  replace_na(list(DOMLevel = 'Not Listed')) %>%
  mutate(Agency = stringr::str_split(GrantAgency, ';')) %>%
  unnest %>%
  mutate(Agency = if_else(Agency %in% c('AMOSO', 'CIHR'), Agency, 'OTHER')) %>%
  group_by(Division, DOMLevel, Agency) %>%
  count() %>%
  spread(DOMLevel, n, fill = 0) %>%
  replace_na(list(JR = 0, SR = 0, MID = 0)) %>%
  ungroup

grant.totals <- 
  grant.reviews %>%
  select(-Division) %>%
  group_by(Agency) %>%
  summarise_all(list(sum)) %>%
  mutate(Division = 'Total')



grant.reviews %>%
  bind_rows(grant.totals) %>%
  select(Division, Agency, JR, MID, SR) %>%
  kable(format = 'latex', align = c('r', 'r', 'c', 'c', 'c')) %>%
  kable_styling(latex_options = c('striped', 'hold_position'))
```